---
title: "pstat 126 project"
format: pdf
editor: visual
---

## PSTAT 126 Final Project

Sruthi Gopanapalli (3468956) and Harika Madiraju (3569092)

**Part 1: Data Description and Descriptive Statistics**

```{r}
data <- read.csv("Diamonds Prices2022.csv")

set.seed(123)
sample_data <- data[sample(nrow(data), 2000), ]
diamonds <- sample_data[,-1] 

```

```{r}
#histograms for continuous variables
hist(diamonds$carat,
main="Histogram of Carat",
xlab="Carat",
col="lightblue",
breaks=30)

hist(diamonds$depth,
main="Histogram of Depth",
xlab="Depth",
col="lightgreen",
breaks=30)

hist(diamonds$table,
main="Histogram of Table",
xlab="Table",
col="lightpink",
breaks=30)

hist(diamonds$price,
main="Histogram of Price",
xlab="Price ($)",
col="lightyellow",
breaks=30)

hist(diamonds$x,
main="Histogram of Length (x)",
xlab="Length (mm)",
col="lightgray",
breaks=30)

hist(diamonds$y,
main="Histogram of Width (y)",
xlab="Width (mm)",
col="lightcoral",
breaks=30)

hist(diamonds$z,
main="Histogram of Depth (z)",
xlab="Depth (mm)",
col="lightcyan",
breaks=30)

#barplots for categorical random variables
barplot(table(diamonds$cut),
       main = "Bar Plot of Diamond Cut",
       xlab = "Cut",
       ylab = "Frequency",
       col = "lightblue")

barplot(table(diamonds$color),
       main = "Bar Plot of Diamond Color",
       xlab = "Color",
       ylab = "Frequency",
       col = "lightgreen")

barplot(table(diamonds$clarity),
       main = "Bar Plot of Diamond Clarity",
       xlab = "Clarity",
       ylab = "Frequency",
       col = "lightpink")

```

(Comment a sentence on distributions for each of the variables)

```{r}
#install.packages("GGally")
#library(GGally)
ggpairs(diamonds[, c("carat","depth","table","price","x","y","z")])

```

```{r, fig.height=10}

model <- lm(price ~ carat + cut + color + clarity + depth + table + x + y + z,
data = diamonds)

summary(model)
plot(model)
```

**Part 2: Simple Linear Regression**

```{r}
model_1 <- lm(price ~ carat, diamonds)
summary(model_1)
plot(diamonds$carat, diamonds$price,
     main="Price vs Carat",
     xlab="Carat",
     ylab="Price (dollars)",
     col="blue")

abline(model_1, col="red")
```

**Interpretations: (update with new numbers)**

The model estimates that for every additional 1 carat increase for the diamond, the price increases an average of \$7,756.44.

The model also predicts that diamonds that are zero carats will have a price of -\$2,256.40.

When performing a hypothesis test on the slope of the model, we have the null hypothesis $H_0$ : $\beta_1 = 0$ and the alternative hypothesis $H_a$ : $\beta_1 \ne 0$. We have a p value of \< 2e-16 which is less than 0.05 so we reject the null hypothesis. This suggests carat weight is a statistically significant predictor of diamond price. The t value is 551.4 which is very large, indicating the slope is approximately 551 standard errors away from 0.

The adjusted $R^2$ is 0.8493 which means that the model explains approximately 84.93% of the variability in diamond prices from just the carat weight. This indicates that there is a very storng relationship between carats and price.

```{r}
confint(model_1)
```

```{r}
predict(model_1,
        newdata = data.frame(carat = 1),
        interval = "confidence",
        level = 0.95)
predict(model_1,
        newdata = data.frame(carat = 1),
        interval = "prediction",
        level = 0.95)
```

Testing the assumptions:

```{r}
# Autocorrelation
Box.test(model_1$fitted.values, type = "Ljung-Box", lag = 10)
# checks if residuals are correlated
# Homoscedasticity
plot(model_1$fitted.values, model_1$residuals)
abline(h = 0, col = "red")
#check if residuals are randomly scattered around 0
# Normality of residuals
qqnorm(model_1$residuals)
qqline(model_1$residuals)
log_linear_model = lm(log(price) ~ carat, data = diamonds)
summary(log_linear_model)
```

**Part 3: Simple Linear Regression continued**
